Explaining NonLinear Classification Decisions with Deep Taylor Decomposition by Montavon et al.

### 2022-06-30 Präsentation/Präsentation_23.06.2020.tex
final version presentation
las changes

\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = w_{ij}$ (gradient of $R_j$)
	\item No domain restriction and for $\tilde{x}_i^{(j)} = x_i - \frac{R_j(x_i)}{\sum_{i'}w_{i'j}^2} w_{ij}$
	\begin{align*}
		R_j(\tilde{x}) &= \max(0, \sum_i (x_i -\frac{R_j(x_i) w_{ij}}{\sum_{i'}w_{i'j}^2}) w_{ij}  + b_j) \\
		&= \max(0, \underbrace{\sum_i (x_i w_{ij} + b_j)}_{=R_j(x_i)} - R_j (x_i) \underbrace{\frac{\sum_{i}w_{ij}^2}{\sum_{i'}w_{i'j}^2}}_{=1} )= 0 
	\end{align*}
	\vspace{-0.1cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

\begin{frame}
\frametitle{$z^b$-rule $\mathcal{X}=\mathcal{B}$}
\vspace{0.4cm}
Given a bounded input space $\mathcal{B} = \{\{x_i\} : \forall_{i=1}^d l_i\leq x_i \leq h_i\} $, with $l_i \leq 0$ and $ h_i \geq 0$ and we search on the segment $(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\},\{x_i\})\subset \mathcal{B}$\\
\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = x_i - l_i 1_{w_{ij}>0} - h_i 1_{w_{ij}<0}$
	\item If $\{x_i\} \in \mathcal{B}$ so is the whole domain and for $w_{ij}^-=\min(0,w_{ij})$, $w_{ij}^+=\max(0,w_{ij})$
	\vspace{-0.25cm}
	\begin{align*}
		R_j(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\}) &= \max(0, \sum_i l_i 1_{w_{ij}>0} w_{ij} + h_i 1_{w_{ij}<0} w_{ij} + b_j)\\
		&=\max(0, \sum_i \underbrace{l_i w_{ij}^+}_{\leq 0} + \underbrace{h_i w_{ij}^-}_{\leq 0} + b_j)= 0
	\end{align*}
	\vspace{-0.75cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

\begin{frame}
\frametitle{Example MNIST: Setting}
\vspace{0.5cm}
Training of a neural network to detect a handwritten digit between 0-3 next to a distracting digit from 4-9 given the following setting:
\begin{itemize}
\item Images of size 28 x 56 pixels and 1568 input neurons $\{x_i\}$
\item One hidden layer with 400 neurons $\{x_j\}$ and one output $x_k$
\item Random initialized weights $\{w_{ij}\}$ and non-negative bias $\{b_j\}$ initialized to zero
\item Training with 300000 iterations of stochastic gradient descent with a batch size of 20 
\end{itemize}
\begin{figure}
\includegraphics[height=2.5cm, width=8cm]{image/fig._1_example}

@@

\begin{frame}
\frametitle{Example MNIST: Heatmaps}
\vspace{0.5cm}
\begin{figure}
\includegraphics[height=2.5cm, width=12cm]{image/example_mnist_heatmap}
\end{figure}

\begin{figure}
\includegraphics[height=3.25cm, width=12cm]{image/example_mnist_consistent}
\vspace{-0.75cm}
\caption{Heatmap and Empirical Results of Consistency}
\end{figure}

@@

\begin{frame}
\frametitle{Deep Networks}
Many problems require very complex deep architectures

@@

 \begin{columns}
          \begin{column}{5cm}
          		$\rightarrow$ Compute $\{v_{ij}, v_{lj},d_j\}$ by minimizing 
			\begin{equation*}
			\min\langle (\hat{R}_k-R_k)^2) \rangle
			\end{equation*}
            \end{column} 
            \begin{column}{6cm}
			\begin{figure}
			\label{fig1}
			\includegraphics[height=3cm, width=6.5cm]{image/min_max_relevance_model}
			\end{figure}
	\end{column}
\end{columns}

@@

$\rightarrow$ The Min-Max relevance model is due to the minimization only approximately consistent
\end{frame}

@@

\item Detection layer
\begin{equation*}
R_i = \sum_j \frac{q_{ij}}{\sum_{i'} q_{i'j}} R_j
\end{equation*}
where $q_{ij}= w_{ij}^2$, $q_{ij}= x_i w_{ij}^+$ or $q_{ij}= x_i w_{ij} - l_i w_{ij}^+ - h_i w_{ij}^-$ for the $w^2$-rule, $z^+$-rule and the $z^b$- rule respectively
\end{itemize}
\vspace{0.1cm}
$\rightarrow$ The training-free relevance model is consistent\\

\vspace{0.5cm}
When using the training-free model for the whole network all but the first layer need to be decomposed using the $z^+$-rule
\end{frame}

>>>>

\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = w_{ij}$ (gradient of $R_j$)
	\item No domain restriction and for $\tilde{x}_i^{(j)} = x_i - \frac{R_j(x_i)}{\sum_{i'}w_{i'j}^2} w_{ij}$
	\begin{align*}
		R_j(\{\tilde{x}_i^{(j)}\}) &= \max(0, \sum_i (x_i -\frac{R_j(x_i) w_{ij}}{\sum_{i'}w_{i'j}^2}) w_{ij}  + b_j) \\
		&= \max(0, \underbrace{\sum_i (x_i w_{ij} + b_j)}_{=R_j(x_i)} - R_j (x_i) \underbrace{\frac{\sum_{i}w_{ij}^2}{\sum_{i'}w_{i'j}^2}}_{=1} )= 0 
	\end{align*}
	\vspace{-0.1cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

\begin{frame}
\frametitle{$z^b$-rule $\mathcal{X}=\mathcal{B}$}
\vspace{0.4cm}
Given a bounded input space $\mathcal{B} = \{\{x_i\} : \forall_{i=1}^d l_i\leq x_i \leq h_i\} $, with $l_i \leq 0$ and $ h_i \geq 0$ and we search on the segment $(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\},\{x_i\})\subset \mathcal{B}$\\
\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = x_i - l_i 1_{w_{ij}>0} - h_i 1_{w_{ij}<0}$
	\item If $\{x_i\} \in \mathcal{B}$ so is the whole domain and for $w_{ij}^-=\min(0,w_{ij})$, $w_{ij}^+=\max(0,w_{ij})$
	\vspace{-0.25cm}
	\begin{align*}
		R_j(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\}) &= \max(0, \sum_i l_i 1_{w_{ij}>0} w_{ij} + h_i 1_{w_{ij}<0} w_{ij} + b_j)\\
		&=\max(0, \sum_i \underbrace{l_i w_{ij}^+}_{\leq 0} + \underbrace{h_i w_{ij}^-}_{\leq 0} +\  b_j)= 0
	\end{align*}
	\vspace{-0.75cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

\begin{frame}
\frametitle{Example MNIST: Setting}
\vspace{0.5cm}
Training of a neural network to detect a handwritten digit between 0-3 next to a distracting digit from 4-9 given the following setting:
\begin{itemize}
\item Images of size 28 x 56 pixels and 1568 input neurons $\{x_i\}$
\item One hidden layer with 400 neurons $\{x_j\}$ and one output $x_k$
\item Random initialized weights $\{w_{ij}\}$ and non-positive bias $\{b_j\}$ initialized to zero
\item Training with 300000 iterations of stochastic gradient descent with a batch size of 20 
\end{itemize}
\begin{figure}
\includegraphics[height=2.5cm, width=8cm]{image/fig._1_example}

@@

\begin{frame}
\frametitle{Example MNIST: Heatmaps}
\vspace{0.5cm}
\begin{figure}
\includegraphics[height=2.5cm, width=12cm]{image/example_mnist_heatmap}
\end{figure}
\pause
\begin{figure}
\includegraphics[height=3.25cm, width=12cm]{image/example_mnist_consistent}
\vspace{-0.75cm}
\caption{Heatmap and Empirical Results of Consistency}
\end{figure}

@@

\begin{frame}
\frametitle{Deep Networks}
Many problems require complex deep architectures

@@

 \begin{columns}
          \begin{column}{5cm}
          		$\rightarrow$ Compute $\{v_{ij}, v_{lj},d_j\}$ by minimizing 
			\begin{equation*}
			\min\langle (\hat{R}_k-R_k)^2 \rangle
			\end{equation*}
            \end{column} 
            \begin{column}{6cm}
			\begin{figure}
			\label{fig1}
			\includegraphics[height=3cm, width=6.5cm]{image/min_max_relevance_model}
			\end{figure}
	\end{column}
\end{columns}

@@

$\rightarrow$ The Min-Max relevance model is due to the minimization only approximately consistent.
\end{frame}

@@

\item Detection layer
\begin{equation*}
R_i = \sum_j \frac{q_{ij}}{\sum_{i'} q_{i'j}} R_j
\end{equation*}
where $q_{ij}= w_{ij}^2$, $q_{ij}= x_i w_{ij}^+$ or $q_{ij}= x_i w_{ij} - l_i w_{ij}^+ - h_i w_{ij}^-$ for the $w^2$-rule, $z^+$-rule and the $z^b$- rule respectively
\end{itemize}
\vspace{0.1cm}
$\rightarrow$ The training-free relevance model is consistent.\\

\vspace{0.5cm}
When using the training-free model for the whole network, all but the first layer need to be decomposed using the $z^+$-rule.
\end{frame}
